---
title: "Homework 3"
author: "Lea Lambdin"
date: 2026-05-28
format: pdf
toc: true
execute:
  warning: false
  message: false
---
```{r}
# load packages
library(tidyverse)
library(here)
library(janitor)

# read in kelp data
kelp <- read_csv(here("data", "temp-kelp.csv"))
```

## Problem 1

### a. An appropriate test

Two appropriate tests for evaluating the relationship between ocean temperature and giant kelp frond elongation rate are a Pearson correlation test and a simple linear regression. Both tests evaluate relationships between two continuous numerical variables, but they differ in purpose. A Pearson correlation measures the strength and direction of the linear relationship between temperature and kelp elongation rate, while a simple linear regression models how kelp elongation rate changes as temperature changes and allows prediction of the response variable from the predictor variable.

### b. Create a visualization

```{r}
# create scatterplot showing relationship between temperature and kelp growth

ggplot(kelp, aes(x = temp_c, y = kelp_elong)) +

  # add points for each observation
  geom_point(
    color = "darkgreen",
    size = 3
  ) +

  # add linear trend line
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "steelblue"
  ) +

  # relabel axes with units
  labs(
    x = "Ocean Temperature (°C)",
    y = "Kelp Frond Elongation Rate (cm/day)",
    title = "Relationship Between Ocean Temperature and Giant Kelp Growth"
  ) +

  # use cleaner theme
  theme_minimal()
```