---
title: "Homework 3"
author: "Lea Lambdin"
date: 2026-05-28
format: pdf
toc: true
execute:
  warning: false
  message: false
---
```{r}
# load packages
library(tidyverse)
library(here)
library(janitor)

# read in kelp data
kelp <- read_csv(here("data", "temp-kelp.csv"))
```

## Problem 1

### a. An appropriate test

Two appropriate tests for evaluating the relationship between ocean temperature and giant kelp frond elongation rate are a Pearson correlation test and a simple linear regression. Both tests evaluate relationships between two continuous numerical variables, but they differ in purpose. A Pearson correlation measures the strength and direction of the linear relationship between temperature and kelp elongation rate, while a simple linear regression models how kelp elongation rate changes as temperature changes and allows prediction of the response variable from the predictor variable.

### b. Create a visualization

```{r}
# create scatterplot showing relationship between temperature and kelp growth

ggplot(kelp, aes(x = temp_c, y = kelp_elong)) +

  # add points for each observation
  geom_point(
    color = "darkgreen",
    size = 3
  ) +

  # add linear trend line
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "steelblue"
  ) +

  # relabel axes with units
  labs(
    x = "Ocean Temperature (°C)",
    y = "Kelp Frond Elongation Rate (cm/day)",
    title = "Relationship Between Ocean Temperature and Giant Kelp Growth"
  ) +

  # use cleaner theme
  theme_minimal()
```

### c. Check assumptions and run your test

#### Assumption checks

```{r}
# create linear model object
kelp_lm <- lm(kelp_elong ~ temp_c, data = kelp)

# plot diagnostic plots for assumptions
par(mfrow = c(2, 2))
plot(kelp_lm)

# perform Shapiro-Wilk test for normality of residuals
shapiro.test(residuals(kelp_lm))
```

The assumptions checked for the linear regression included linearity, normality of residuals, and homoscedasticity. I checked these assumptions using diagnostic plots from the linear model and a Shapiro-Wilk test on the residuals. Based on the diagnostic plots and test results, the assumptions appeared reasonably met for this dataset.

#### Run the test

```{r}
# run simple linear regression
summary(kelp_lm)

# run Pearson correlation test
cor.test(kelp$temp_c, kelp$kelp_elong)
```

### d. Results communication

To evaluate the strength of the relationship between ocean temperature and giant kelp frond elongation rate, I used a simple linear regression. I selected this test because both variables were continuous numerical variables, and I wanted to determine how changes in temperature were associated with changes in kelp elongation rate.

The results showed a significant negative relationship between ocean temperature and giant kelp frond elongation rate (F = 26.93, p < 0.001, R² = 0.473). As ocean temperature increased, kelp elongation rate tended to decrease. The Pearson correlation test also indicated a moderately strong negative correlation between the variables (r = -0.688, p < 0.001).

### e. Test implications

These results suggest that warmer ocean temperatures may negatively impact giant kelp growth rates. Since giant kelp forests are important marine ecosystems that provide habitat and food for many species, reduced growth under warmer conditions could have broader ecological consequences. This information could help researchers and conservation teams monitor vulnerable kelp forests and prioritize areas experiencing increasing ocean temperatures.

### f. Double check your own work

The Pearson correlation test and the simple linear regression led to the same overall conclusion about the null hypothesis. Both tests showed a significant negative relationship between ocean temperature and giant kelp frond elongation rate, with p-values less than 0.001. The Pearson correlation quantified the strength and direction of the linear relationship between the two variables, while the linear regression modeled how kelp elongation rate changed as temperature increased. Together, the tests supported the interpretation that higher temperatures were associated with lower kelp growth rates.